Collision rates of planetesimals near mean-motion resonances

### 2021-01-25 refereeResponse.txt

    This section is meant to focus solely on the dynamical effects that the planet has on the full-sized planetesimal population, in order to motivate the initial conditions used in the next section, and also to explain the planetesimal collision profiles shown in section 4.1. We are purposefully trying to keep discussion of radial drift and collisional debris as a separate, encapsulated part of the paper (section 5) in the new version.

@@

20) pg.3 rcol P.2 l.34 - "The planetesimal disc extends from 2.2 to 3.8 AU..." - Why not consider the outer disk too? External resonances can be very strong and capture is more likely there.

    This is a good point. Our work was mainly motivated by the non-result seen by Richardson et al. 2000, which is why we focus on interior MMRs here. We have also added a brief discussion highlighting the importance of studying exterior resonances in future work in the summary and discussion near the end of the paper.

21) pg.3 rcol P.2 l.41 - This is a minor comment. I am just curious. Why not using alpha = -1 or -0.9 as it seems more common for protoplanetary disks? -1.5 is an approximation for the Minimum Mass Solar Nebula based on the current solar system planets. Observations of proptoplanetary disks sugggest alpha = -1 or -0.9.

    See 19) and 20). alpha = -1,5 has been a standard benchmark in the field for quite a while and is we do not think it's unreasonable to use. Furthermore, there is evidence to suggest that the solid surface density profile close to the star is steeper than alpha = -1 (see Chiang + Laughlin 2013, and Dai, F 2020). The region of the protoplanetary disk that we are modeling is not well-constrained by either exoplanet studies or mm and sub-mm observations.

@@

22) pg.3 rcol P.2 l.44 - "Each simulation is evolved for 5,000 years" - Could you, please, add discussion regarding simulation time, as well as, define what is a short and long time. For me 5000 yr is, in principle, too short (see my concern #4).

    See our response to concern #4. Some text has been added to the paper (at the beginning of section 4) to elaborate on this.

23) pg.3 rcol P.3 l.44-46 - "The dynamical effects of secular forcing by the giant planet, along with the effects of viscous stirring and gas drag on the planetesimals, are built into the initial conditions." - From the text, It is not clear whether the authors are considering the gas disk phase or not?

    See 8). We are considering the gas disk phase here. This shouldn't matter too much in terms of the motion and collisions of the planetesimals, although the presence of the gas is very important to construct the dust profiles. More text has been added to this section to explain this more clearly.

@@

27) pg.4 rcol P.1 l.41 - "... T_lib ~ 1,000 - 2,000 years..." - So the integration time is about a factor of 2-4 T_lib. Is this enough to take conclusions? Please, see my concern and add discussion.

    In our case, the large particle count allows us to look at the 'ensemble average' of the dynamical behavior near resonance. Even though individual planetesimals only go through a few librational oscillations, the phase trajectories that they follow are well populated with particles. This allows us to make conclusions about the dynamics without needing a longer baseline from which to take time averages. Had we been following only a few bodies, this would not have been the case. Some additional text has been added to this paragraph to explain this more clearly.

28) pg.4 rcol P.2 l.52-53 - "Resonances with the perturbing planet are visible as nonaxisymmetric gaps." - I am afraid that this is only because none of your particles are drifting, which would locally increase the collision rates once small planetesimals fragments that should be drifting away start to (maybe unrealistically) pile up within or near MMR, thus quickly collisionally evolving to dust. I wonder if this would ever happen if gas drag was to play? Please, see my concern and add discussion.

    This is referring to the positions of the planetesimals in this figure, and not to any kind of debris. Even if the smaller fragments do drift significantly, the nonaxisymmetric gaps in the planetesimal distribution should still be present. Of course, this would wash out all of the interesting features in the dust emission profile. This sentence has been modified to clarify that we are specifically talking about the structure seen in figure 1.

@@

29) pg.4 rcol P.2 l.56-59 - "... in Richardson et al. (2000) (see bottom panel of figure 3) and Tabeshian & Wiegert (2016) (figure 3)." - add figure 3 in those works. Fig 3 in this work does not have bottom panels.

    The text now specifies the subfigures we are referring to (3c and 3f in Richarson, 3a in Tabeshian & Weigert)

30) Figure 2 - I wonder how (or if) particle drifting due to gas-drag would change this figure? Also, the 8:3 and 7:2 MMRs are irrelevant here! Why not showing 7:3 MMR that is discussed in the main text?

    As we now calculate in section 2.1, the libration timescale and the secular forcing timescale are much shorter than the stopping time due to gas for the planetesimals, so this figure should not look different had we modeled gas drag in the simulations. A typo in the text has been corrected: '7:3' was supposed to say '8:3'.

@@

35) pg.6 lcol P.1 l.33 - "cylindrical distance" - please, define before start using it.

    A sentence has been added to the text to clarify what cylindrical distance means.

36) pg.6 rcol P.2 l.47-54 - "This also suggests... radial collision profile." - I don't know if I agree with authors' conclusions that MMRs are not that important, and that most collisions occur with objects out of resonance. It looks for me that the features seem in Fig 6 (similar to Fig 4) are caused by objects in MMRs colliding at their pericentric and apocentric distances within the eccentric orbits. In my understanding, if authors' arguments are correct, MMRs should basically not matter, and Fig 5 would look similar if no planet was present. If that is the case, how can this work serve to constrain the mass and eccentricity of a perturbing planet?

    As shown in figure 6 (fig 5 in the original version), the bump and dip features at the center of the MMRs are still present, even if we ignore particles whose semimajor axis falls within the libration width of the resonances. What this implies is that the edges of the resonances are important. As we approach a resonance from the outside, the critical angle \phi becomes near-stationary (using the pendulum analogy, this corresponds to a bob in the vertical, upside-down position). Where the critical angle is near-stationary, the pericenter orientations of the planetesimals experience a torque which causes their \varpi to follow the giant planet in its orbit. This is what we propose is responsible for the coherent structure seen in the right-hand panels of figures 7 and 8 (6 and 7 in the original paper), along with the structure seen in the collision profiles.

@@

    The paragraph referenced now clarifies that we are specifically referring to collisions between bodies inside of the libration width of the resonance. An additional sentence has been added at the end of the following paragraph (which provides an explanation for the nonaxisymmetric gaps seen in fig 1) to tie this back to figure 6.

37) pg.6 rcol P.2 l.38-56 - "As mentioned in section 2, ... appears to 'follow' the planet." - This paragraph is super confuse. I don't understand this. Why not simply say this is caused by the secular forcing in eccentricity? This would save 3 paragraphs in the paper and would be much more clear.

    See 36). If you look at figures 1, 7 and 8, (1, 6 and 7 in the original paper) the nonaxisymmetric gap structure in the disk near the resonances is always aligned with the planet's orbital phase. It is also apparent from figure 1 that the gaps are not aligned with the planet's longitude of pericenter, which should be the case if the structure were due to secular forcing in eccentricity. An explanation similar to the referee's was given by Tabeshian & Weigert 2016, which, as we point out, does not explain why the gaps are misaligned from the planet's longitude of pericenter.

@@

38) Figs 4 and 5 - Why is the relative # larger than 1? Relative to what?

    These are showing the probability density of the collision rates as a function of semimajor axis (fig 4) and distance from the star (fig 5). The curves on these plots are normalized such that the area underneath = 1. For the orange curve in figure 5, the normalization factor is scaled by the ratio between the total number of collisions in each sample. Some text has been added alongside the discussion of figures 4 and 5 to explain this.

@@

39) General note about the figures - Not that I want to be repetitive, but I am afraid that if small planetesimals are allowed to drift the bumps and dips from these figures may change as the small planetesimals would likely collide everywhere in radial distances. This is because internal MMRs are not strong enough to perturb these small fast-drifting planetesimals. In my opinion, without considering small planetesimals drifting due to gas drag, Figs 6 and 7 (especially) represent a planetesimal collisional profile, but not necessarily a dust profile. Please, see my concerns and add discussion.

    See our response to concern 4)

40) pg.10 lcol P.4 l.42-44 - "Many different parameterizations are possible, depending on optical depth, grain size, disc temperature, etc.,..." - But couldn't this change all of your conclusions? Please, see my concerns regarding planetesimal sizes and add discussion.

    To clarify, these parameters will alter the overall brightness profile of the dust, but should not affect the features due to the resonances. Some additional text has been added to this section to explain this.

@@

41) Last paragraph of section 5 - "In order to... cases shown above" - This is a big IF, especially if what the authors are measuring here is the planetesimal collisional profile rather than the dust profile, or if the dust profile would be even shallower as a consequence of aerodynamic drag acting on small fragments of collisionally evolved planetesimals.

    This is certainly a big 'if', although we have shown with a back-of-the-envelope calculation in section 5 that the connection between the collision profile and the dust profile is not completely unreasonable. We have added some additional text to the end of section 5 to remind the reader that the dust profile is built on somewhat tenuous assumptions.

@@

42) pg.12 rcol P.1 l.8-24 - "Near the interior 2:1... both of these quantities." - This is a clear way to set the explanation. Well done! I hope to see more of that in the revised version. I think a lot of the explanations in the main text should be substituted by something clear and concise as this! Make it simple!

    :-)

43) pg.12 rcol P.3 l.45-48 - "Given that the solid surface density diminishes with distance (r^-1.5 for the MMSN (Hayashi 1981)), exterior resonances may not be quite as useful for this type of measurement, however.)" - Not sure I agree. Exterior MMRs can more easily capture inward drifting objects (e.g. Weidenschilling & Davis 1985; Patterson 1987; Beauge & Ferraz-Mello 1993; Malhotra 1993; Gomes 1995; Jiang & Yeh 2004; Batigyn & Laughlin 2015; Storch & Batygin 2019), increasing their local collision rates (Deienno et al. 2020), thus potentially generating very strong dusty features.

    This is absolutely true if we are going to consider radial drift. Given that an expanded discussion of radial drift is now an important part of this paper, we should be more careful not to dismiss the importance of exterior MMRs. Here, we have added a sentence mentioning that exterior MMRs might be important if radial drift is modeled.

@@

44) pg.12 rcol P.4 l.50-52 - "As we have shown, the bump vs dip feature near the 2:1 MMR is marginally detectable with ALMA for the nearest protoplanetary discs." - Please, see my concern and add discussion.

    This is addressed in a few other places in the new version of the paper, but we have added a sentence here reminding the reader of the assumption that goes into constucting the dust profile.

My apologies for being so long. I do think this work has potential, but it does need substantial clarifications.

	No need to apologize. Thank you very much for being so thorough!

	References used in response:

	- Chiang, E. & Laughlin, G., 2013, MNRAS, 431, 3444C
	- Dai, F. et al., 2020, AJ, 159, 247D
